Cell Decompositions for Rank Two Quiver Grassmannians

### 2018-02-27 Grassmannian_cells.tex
Prevent line breaks in references

  As there exist representations with trivial endomorphism ring, the dimension vectors $d(m,r)$ are Schur roots.
  It follows that the set of indecomposable representations with trivial endomorphism ring forms a dense open subset of $R_{d(m,r)}(K(n))$, see for example \cite[Theorem 2.2]{sch}.
  This shows the last claim.
\end{proof}
\begin{remark}
  There is a more elegant way to prove the first part of Proposition \ref{pro:indecomposables} using the notion of stability and moduli spaces.
  Actually, fixing the standard stability induced by the linear form $\Theta:\ZZ^{Q_0}\to\ZZ$ defined by $\Theta(d)=d_2$, it can be shown that all indecomposables are stable and that the moduli space of stable representations is in fact $\Gr_l(\CC^n)$.
  We opted for the proof above because the notion of stability would only be used at this point and we wanted to keep the exposition as simple as possible. 
\end{remark}

\begin{lemma}
  \label{le:basic homological properties}
  For $V\in\Gr(H_m)$, $m\ge1$, we have $\Hom(P_{m+1}^V,P_\ell)=0=\Ext(P_\ell,P_{m+1}^V)$ for all $\ell\ge1$.
\end{lemma}
\begin{proof}
  Recall that $\Hom(X,P_\ell)\ne0$ (resp. $\Ext(P_\ell,X)\ne0$) for some indecomposable representation $X$ and some $\ell\ge1$ implies that $X$ is preprojective of the form $P_r$ with $1\le r\le\ell$ (resp. $1\le r\le\ell-2$).
  However, $P_{m+1}^V$ is indecomposable by Proposition \ref{pro:indecomposables} and it cannot be preprojective as it is not rigid by Lemma~\ref{le:unique morphisms}.
  %Note that $\udim P_{m+1}^V <\udim P_{m+1}$, but $\udim P_r\geq \udim P_{m+1}$ if $\Hom(P_{m+1},P_r)\neq 0$.
\end{proof}

@@

  For the $m=3$ case, we note that $\tau\tilde P_4^I=\tilde P_2^I$ by Lemma~\ref{le:truncated tau}.
  Then the claimed structure $\tilde P_3(I,j)$ of the kernel and the surjectivity of the map $\tilde P_{3,j}\to\tau\tilde P_4^I$ are immediate from the explicit descriptions of $\tilde P_{3,j}$ and $\tilde P_2^I$ in Example~\ref{ex:lifted preprojectives} and Example~\ref{ex:truncated lifts} respectively.
  The general cases for parts (1) and (2) both then follow using the reflection recursions \eqref{eq:recursive covers} taking into account Remark \ref{rem:reflection recursion}.(2).
\end{proof}

@@

\subsection{Torus Actions and the Bia\l{}ynicki-Birula Decomposition}
The aim of Section \ref{torusaction} is to define a torus action on quiver Grassmannians which can be used to simplify the calculation of homological invariants in general.
If the quiver Grassmannian is smooth, which is for instance the case for exceptional representations by \cite{cr}, it can also be used to stratify the quiver Grassmannians using the results of Bia\l{}ynicki-Birula.
More specifically, let $X$ be a smooth projective variety with a $\C^\ast$-action.
For a connected component of the fixed point set $C\subset X^{\C^\ast}$, we define its attracting set as
\[\Att(C):=\{y\in X\mid \lim_{t\to 0}t.y\in C\}.\]
The following result of Bia\l{}ynicki-Birula relates the geometry of $X$ to the geometry of its torus fixed points.
\begin{theorem}
  \label{thm:bb}
  \cite[Section 4]{bb}
  Each attracting set $\Att(C)$ is a locally closed $\CC^*$-invariant subvariety of $X$ and the natural map $\Att(C)\to C$ is a vector bundle.
  Moreover, assuming $X^{\C^\ast}=\coprod_{i=1}^r C_i$ is a decomposition of the fixed point set of $X$ into finitely many connected components, we have $X=\coprod_{i=1}^r \Att(C_i)$.
  In particular, we have an equality of Euler characteristics $\chi(X)=\chi(X^{\C^\ast})$.
\end{theorem}

@@

%Any subrepresentation $E\subset P_{m+1}^V$ corresponds by taking pullbacks to a chain of inclusions $P_m\otimes V\subset\tilde E\subset P_{m+1}$.
Write $\Ind(K(n),d)$ for the set of isomorphism classes of indecomposable representations of $K(n)$ of dimension $d$ for $d\in\NN^{K(n)_0}$. As the $\GL_n(\CC)$-action commutes with the natural base change action, we can define a $\GL_n(\CC)$-action on $\Ind(K(n),d)$.
\begin{proposition} \label{indecomposables}
  Let $m\geq 1$ and $d(m,r)=\dim P_{m+1}-r\dim P_m$ with $0\leq r\leq n-1$. The action of $\GL_n(\CC)$ is transitive on $\mathrm{Ind}(K(n),d(m,r))$.
\end{proposition}
\begin{proof}
  For $m=1$, the action of $\GL_n(\CC)$ on $\Gr_d(\CC^n)$ is transitive which shows that it is transitive on $\Ind_{d(1,r)}(K(n))$.
  As the reflection functors preserve indecomposability and isomorphism classes, by Lemma \ref{compatibleGSigma}, for every $g\in\GL_n(\CC)$, we obtain a commutative diagram
  \[\xymatrix{\Ind(K(n),d(m,r))\ar[r]^{\Sigma_2}\ar[d]^{g^{-T}}&\Ind(K(n),d(m+1,r))\ar[d]^g\\
    \Ind(K(n),d(m,r))\ar[r]^{\Sigma_2}&\Ind(K(n),d(m+1,r))}.\]
  As the $\GL_n(\CC)$-action is transitive on the left hand side, this implies that it is also transitive on the right hand side.
\end{proof}

@@

  Proposition \ref{pro:indecomposables} and \ref{indecomposables} imply that all quiver Grassmannians $\Gr_\bfe(M)$ for a fixed $\bfe\in\NN^{Q_0}$ where $M\in\Ind(K(n),\udim P_{m+1}^V)$ are isomorphic.
  With other words we use that all indecomposables are truncated preprojectives.

@@

  The second statement follows directly from Proposition \ref{quotient}.
\end{proof}

@@

By induction hypothesis the quiver Grassmannians $\Gr^{\tilde Q}_\tbff( \tilde P_{m-1,j_l})$ have cell decompositions which are compatible with $\pi^{\bf i}_m$ and, moreover, the quiver Grassmannians $\Gr^{\tilde Q}_\tbfg(\tilde P_{m}^{\uj})$ have a cell decompositions which are compatible with $\pi^{\bf j}_{m,\uj}$. The second part of Proposition \ref{fibers} says that the fibers are empty if and only if $\pi^{\bf i}_m=0$ and $\tbfg=\udim \tilde P_{m+1}^{(k)}$. This yields a collection of cells over which the fibers are empty. Note that $\Gr^{\tilde Q}_\tbfg(\tilde P_{m}^{\uj})$ is just a point in this case.

If $\tbfg\neq \udim\tilde P_{m+1}^{(k)}$, by Proposition \ref{fibers}, the fiber dimension only depends on the dimension vectors $\tbff, \tbfg$ which means that they are constant over the cells of $\Gr^{\tilde Q}_\tbff(P_{m-1,j_l})\times \Gr^{\tilde Q}_\tbfg(\tilde P_{m}^{\uj})$. In particular, they are constant over cells whose subrepresentations $U\in \Gr^{\tilde Q}_\tbff(P_{m-1,i_k})$ satisfy $\pi_m^{\bf i}(U)\neq 0$, but also over those whose subrepresentations $U\in \Gr^{\tilde Q}_\tbff(P_{m-1,i_k})$ satisfy $\pi_m^{\bf i}(U)= 0$. By Theorem \ref{vb}, it follows that $\Gr^{\tilde Q}_\tbfe(\tilde P_{m}^{\uj'})$ has a cell decomposition.

Thus it remains to show that this cell decomposition of $\Gr^{\tilde Q}_\tbfe(\tilde P_{m}^{\uj'})$ is compatible with $\pi_{m,\uj'}^{\bf j}$ where ${\bf j}=(\underline{j_1},\underline{j_2},\ldots,\underline{j_r})$ with $\underline{j_1}=(\uj',j_{1,l},\ldots,j_{1,n-1})\in I_{n-1}$ and $j_{1,l}=j_l$. Recall that we have $\pi_{m,\uj'}^{\bf j}=\pi_{m,\uj}^{\bf j}\circ\pi_{m,\uj'}^{j_l}$ by Lemma \ref{le:truncated quotients}.(2).

@@

By Lemma \ref{directsums} and induction hypothesis every quiver Grassmannian on the right hand side has a cell decomposition. The second part of the Proposition says that the fibers are empty if and only if $\tbfg=0$ and $\tbfh=\udim \tilde P_{m+1}^{(k)}$. In all other cases, the fiber dimension only depend on the dimension vectors $\tbff, \tbfg,\tbfh$ which means that they are constant over $\Gr^{\tilde Q}_\tbff( P(m,k))\times \Gr^{\tilde Q}_\tbfg(\tau\tilde P_{m+1}^{(k)})\times \Gr^{\tilde Q}_\tbfh(\tilde P_{m+1}^{(k)})$\footnote{actually this should already say that this is a vector bundle as the base space is smooth (all reps are exceptional), apply Lemma \ref{vb}?}. Analogously to Lemma \ref{directsums}, we can conclude that this induces trivial bundles over every affine cell. In total, this yields a cell decomposition of the quiver Grassmannian $\Gr^{\tilde Q}_\tbfe(\tilde P_{m+1}^{(k-1)})$.
\end{comment}
\end{proof}

@@

\begin{theorem}\label{celldec}The following holds:
\begin{enumerate}
\item Every quiver Grassmannian of any indecomposable preprojective or preinjective representation of $K(n)$ and $\widetilde{K(n)}$ has a cell decomposition into affine spaces.
\item Let $X\in\rep K(n)$ be an indecomposable representation of dimension $(d,e)$ or $(e,d)$ where $(d,e)=\udim P_m-r\udim P_{m-1}$ for $m\geq 2$ and $\leq 0\leq r\leq n-1$. Then every quiver Grassmannian $\Gr_{\bfe}(X)$ has a cell decomposition into affine spaces.
\end{enumerate}
\end{theorem}
\begin{proof}
The first part follows by combining the results of Section \ref{torusaction} and Theorems \ref{thm:bb} and \ref{cellscover} taking into account that every preprojective representation of $\widetilde{ K(n)}$ is a lift of a preprojective representation which can be proven using the same argument as in the proof of Lemma \ref{le:lifts of transjectives}. The dual version for preinjective representations follows as $\Gr_{\bfe}(P_m)\cong \Gr_{\udim P_m-\bfe}(I_m)$ and $\Gr_{\tbfe}(\tilde P_m)\cong\Gr_{\udim \tilde P_m-\tbfe}(\tilde I_m)$ respectively.

The second part follows in the same way, additionally, taking Theorem \ref{thm:truncpp} into account.
\end{proof}

@@

By Theorem \ref{celldec}, the claim is true for all quiver Grassmannians attached to $P_{j_i}$ or $I_{k_i}$. Consider the short exact sequence
\[\ses{P_{j_{r'+1}}}{P(r'+1)}{P(r')}.\]
By induction, we can assume that all quiver Grassmannians attached to the two outer terms have a cell decomposition into affine spaces. Consider the Caldero-Chapoton map
\[\Psi_\bfe:\Gr_\bfe(P(r'+1))\to\bigsqcup_{\bff+\bfg=\bfe} \Gr_\bff(P_{j_{r'+1}})\times \Gr_\bfg(P(r')).\]
The results of \cite[Section 3]{cc} show that $\Psi_{\tbfe}^{-1}(U,V)=\AA^{\dim\Hom(V,P_{j_{r'+1}}/U)}$  for all $(U,V)\in \Gr_\bff(P_{j_{r'+1}})\times \Gr_\bfg(P(r'))$, in particular the fibre is never empty. Now every subrepresentation $V$ of $P(r')$ is isomorphic to direct sum of preprojective representations such that for each direct summand $P_l$ we have $l\leq j_{r'}$. Moreover, the quotient $P_{j_{r'+1}}/U$ is not projective if $U\neq 0$ and equal to $P_{j_{r'+1}}$ otherwise. Together this yields $\Ext(V,P_{j_{r'+1}}/U)=0$ and thus
$$\dim\Hom(V,P_{j_{r'+1}}/U)=\langle V, P_{j_{r'+1}}/U\rangle$$
for all $(U,V)\in \Gr_\bff(P_{j_{r'+1}})\times \Gr_\bfg(P(r'))$. By Theorem \ref{vb}, this already shows that $\Gr_{\bfe}(P(r'))$ has a cell decomposition into affine spaces for every $1\leq r'\leq r$. By duality, the same is true for $\Gr_{\bfe}(I(s))$.

@@

As the $F$-polynomial of truncated preprojective representations only depend on the dimension vector, we may denote them by $F_{\udim P_m{(k)}}$. The description of the non-empty fibers in Proposition \ref{fibers} together with Corollary \ref{fpoly} and Theorem \ref{celldec} yield the following:
\begin{corollary}
For $m\geq 2$ and $0\leq k\leq n-2$, we have 
$$F_{\udim P_m^{(k)}}=F_{\udim P_m^{(k+1)}}F_{\udim P_{m-1}}-x^{\udim P_{m}^{(k)}}F_{\udim P_{m-2}^{(k)}}.$$
\end{corollary}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Combinatorial description of non-empty cells}
\noindent In this section, we provide two combinatorial descriptions of the non-empty cells in the quiver Grassmannians of (truncated) preprojective representations of $K(n)$.
The first is quiver theoretic and follows directly from the recursive construction of the cell decomposition in Section \ref{sec:fibrations}.
The second is the notion of compatible pairs in a maximal Dyck path arising in the computation of rank 2 cluster variables \cite{llz}.
We give a bijection between these which provides a partial answer to a question of the first author \cite{rupel}.

@@

%The aim is to associate a $2$-quiver with every preprojective representation $P_m$ and with their quotients $P_m^{(k)}$ defined in section \ref{QG}.
%These $2$-quivers are related to certain coefficient quivers of these representations.
%The upshot is that the non-empty cells of $\Gr_e(P_m)$ and $\Gr_e(P_m^{(k)})$ are in one-to-one correspondence with the strong successor closed subsets $\beta$ of type $e$.
%Hence the corresponding Euler characteristic can be computed when counting these subsets.
%It will turn out that a feature of this description is that it does not take care of the coloring of the different arrows induced by the arrows of $K(n)$.

%\begin{remark}\footnote{Insert a remark that we do not work on the universal cover} 
%  Lemma \ref{AR} can be generalized to the cases $m\leq 2$.
%  The only technical problem is that $\tau \tilde P^{(k)}_{m+1}$ is no quotient of $\tilde P_{m,k}$ and we have to replace it by the image of the unique morphism to $\tau \tilde P^{(k)}_{m+1}$ which induces the exact sequence. 
%\end{remark}
In the following, we frequently use the notation and conventions of Section~\ref{Lifting}.
The next step is to associate a $2$-quiver $\cQ_{m+1}^{[k]}$ with the lifted truncated preprojective representations 
\[\tilde P_{m+1}^{[k]}:=
  \begin{cases}
    \tilde P_{m+1}^{\{n,n-1,\ldots,n-k+1\}} & \text{ if $m$ is odd;}\\
    \tilde P_{m+1}^{\{1,2,\ldots,k\}} & \text{ if $m$ is even;}
  \end{cases}\] 
for $m\geq 1$ and $0\leq k\leq n-1$.
When $k=0$ above, we take $\tilde P_{m+1}^{[0]}=\tilde P_{m+1}$.
We aim to construct $\cQ_{m+1}^{[k]}$ so that the strong successor closed subsets are in one-to-one correspondence with cells of quiver Grassmannians of $\tilde P_{m+1}^{[k]}$.

@@

\item  For $m\geq 3$, recall that we have $\dim\Hom(P_{m-1},\tau P_m^{[k]})=\dim\Ext(P_m^{[k]},P_{m-1})$ by the Auslander-Reiten formula. This induces isomorphisms
\begin{align}
  \label{eq:homext} 
  \bigoplus_{i=k+1}^n\Hom(\alpha_{i}^{(-1)^{m}}.\tilde P_{m-1},\tau\tilde P_{m}^{[k]})\cong\bigoplus_{i=k+1}^n\Ext(\tilde P_m^{[k]},\alpha_{i}^{(-1)^{m}}. \tilde P_{m-1})\cong \Ext(P_m^{[k]},P_{m-1}).
\end{align}
This means that we can choose a basis of $\Ext(P_m^{[k]},P_{m-1})$ which is compatible with the lifts to the universal covering and which gives a one-to-one correspondence between the basis elements of $\Ext(P_m^{[k]},P_{m-1})$ and $\Hom(P_{m-1},\tau P_m^{[k]})$. 
In view of Corollary \ref{cor:perpendicular}, we have also an isomorphism $$\Ext(\tilde P_m^{[k]},\tilde P_{m-1,k})\cong\Ext(\tilde P_m^{[k]},\tau\tilde P_m^{[k]}).$$
These isomorphisms will also be present in the structure of the $2$-quivers $\cQ_m^{[k]}$. Finally note that there exists a short exact sequence
\[\ses{\Hom(P_{m-1},P_{m-1})}{\Ext(P_m^{[k]},P_{m-1})}{\Ext(P_m^{[k+1]},P_{m-1})}.\]
Thus a basis of $\Ext(P_m^{[k]},P_{m-1})$ can be obtained by taking the last $k$ basis elements of a basis of $\Ext(P_m^{[n-1]},P_{m-1})$.
\end{enumerate}
\end{remark}

>>>>

  As there exist representations with trivial endomorphism ring, the dimension vectors $d(m,r)$ are Schur roots.
  It follows that the set of indecomposable representations with trivial endomorphism ring forms a dense open subset of $R_{d(m,r)}(K(n))$, see for example \cite[Theorem 2.2]{sch}.
  This shows the last claim.
\end{proof}
\begin{remark}
  There is a more elegant way to prove the first part of Proposition~\ref{pro:indecomposables} using the notion of stability and moduli spaces.
  Actually, fixing the standard stability induced by the linear form $\Theta:\ZZ^{Q_0}\to\ZZ$ defined by $\Theta(d)=d_2$, it can be shown that all indecomposables are stable and that the moduli space of stable representations is in fact $\Gr_l(\CC^n)$.
  We opted for the proof above because the notion of stability would only be used at this point and we wanted to keep the exposition as simple as possible. 
\end{remark}

\begin{lemma}
  \label{le:basic homological properties}
  For $V\in\Gr(H_m)$, $m\ge1$, we have $\Hom(P_{m+1}^V,P_\ell)=0=\Ext(P_\ell,P_{m+1}^V)$ for all $\ell\ge1$.
\end{lemma}
\begin{proof}
  Recall that $\Hom(X,P_\ell)\ne0$ (resp. $\Ext(P_\ell,X)\ne0$) for some indecomposable representation $X$ and some $\ell\ge1$ implies that $X$ is preprojective of the form $P_r$ with $1\le r\le\ell$ (resp. $1\le r\le\ell-2$).
  However, $P_{m+1}^V$ is indecomposable by Proposition~\ref{pro:indecomposables} and it cannot be preprojective as it is not rigid by Lemma~\ref{le:unique morphisms}.
  %Note that $\udim P_{m+1}^V <\udim P_{m+1}$, but $\udim P_r\geq \udim P_{m+1}$ if $\Hom(P_{m+1},P_r)\neq 0$.
\end{proof}

@@

  For the $m=3$ case, we note that $\tau\tilde P_4^I=\tilde P_2^I$ by Lemma~\ref{le:truncated tau}.
  Then the claimed structure $\tilde P_3(I,j)$ of the kernel and the surjectivity of the map $\tilde P_{3,j}\to\tau\tilde P_4^I$ are immediate from the explicit descriptions of $\tilde P_{3,j}$ and $\tilde P_2^I$ in Example~\ref{ex:lifted preprojectives} and Example~\ref{ex:truncated lifts} respectively.
  The general cases for parts (1) and (2) both then follow using the reflection recursions \eqref{eq:recursive covers} taking into account Remark~\ref{rem:reflection recursion}.(2).
\end{proof}

@@

\subsection{Torus Actions and the Bia\l{}ynicki-Birula Decomposition}
The aim of Section~\ref{torusaction} is to define a torus action on quiver Grassmannians which can be used to simplify the calculation of homological invariants in general.
If the quiver Grassmannian is smooth, which is for instance the case for exceptional representations by \cite{cr}, it can also be used to stratify the quiver Grassmannians using the results of Bia\l{}ynicki-Birula.
More specifically, let $X$ be a smooth projective variety with a $\C^\ast$-action.
For a connected component of the fixed point set $C\subset X^{\C^\ast}$, we define its attracting set as
\[\Att(C):=\{y\in X\mid \lim_{t\to 0}t.y\in C\}.\]
The following result of Bia\l{}ynicki-Birula relates the geometry of $X$ to the geometry of its torus fixed points.
\begin{theorem}
  \label{thm:bb}
  \cite[Section 4]{bb}
  Each attracting set $\Att(C)$ is a locally closed $\CC^*$-invariant subvariety of $X$ and the natural map $\Att(C)\to C$ is a vector bundle.
  Moreover, assuming $X^{\C^\ast}=\coprod_{i=1}^r C_i$ is a decomposition of the fixed point set of $X$ into finitely many connected components, we have $X=\coprod_{i=1}^r \Att(C_i)$.
  In particular, we have an equality of Euler characteristics $\chi(X)=\chi(X^{\C^\ast})$.
\end{theorem}

@@

%Any subrepresentation $E\subset P_{m+1}^V$ corresponds by taking pullbacks to a chain of inclusions $P_m\otimes V\subset\tilde E\subset P_{m+1}$.
Write $\Ind(K(n),d)$ for the set of isomorphism classes of indecomposable representations of $K(n)$ of dimension $d$ for $d\in\NN^{K(n)_0}$. As the $\GL_n(\CC)$-action commutes with the natural base change action, we can define a $\GL_n(\CC)$-action on $\Ind(K(n),d)$.
\begin{proposition} \label{indecomposables}
  Let $m\geq 1$ and $d(m,r)=\dim P_{m+1}-r\dim P_m$ with $0\leq r\leq n-1$. The action of $\GL_n(\CC)$ is transitive on $\mathrm{Ind}(K(n),d(m,r))$.
\end{proposition}
\begin{proof}
  For $m=1$, the action of $\GL_n(\CC)$ on $\Gr_d(\CC^n)$ is transitive which shows that it is transitive on $\Ind_{d(1,r)}(K(n))$.
  As the reflection functors preserve indecomposability and isomorphism classes, by Lemma~\ref{compatibleGSigma}, for every $g\in\GL_n(\CC)$, we obtain a commutative diagram
  \[\xymatrix{\Ind(K(n),d(m,r))\ar[r]^{\Sigma_2}\ar[d]^{g^{-T}}&\Ind(K(n),d(m+1,r))\ar[d]^g\\
    \Ind(K(n),d(m,r))\ar[r]^{\Sigma_2}&\Ind(K(n),d(m+1,r))}.\]
  As the $\GL_n(\CC)$-action is transitive on the left hand side, this implies that it is also transitive on the right hand side.
\end{proof}

@@

  Proposition~\ref{pro:indecomposables} and~\ref{indecomposables} imply that all quiver Grassmannians $\Gr_\bfe(M)$ for a fixed $\bfe\in\NN^{Q_0}$ where $M\in\Ind(K(n),\udim P_{m+1}^V)$ are isomorphic.
  With other words we use that all indecomposables are truncated preprojectives.

@@

  The second statement follows directly from Proposition~\ref{quotient}.
\end{proof}

@@

By induction hypothesis the quiver Grassmannians $\Gr^{\tilde Q}_\tbff( \tilde P_{m-1,j_l})$ have cell decompositions which are compatible with $\pi^{\bf i}_m$ and, moreover, the quiver Grassmannians $\Gr^{\tilde Q}_\tbfg(\tilde P_{m}^{\uj})$ have a cell decompositions which are compatible with $\pi^{\bf j}_{m,\uj}$. The second part of Proposition~\ref{fibers} says that the fibers are empty if and only if $\pi^{\bf i}_m=0$ and $\tbfg=\udim \tilde P_{m+1}^{(k)}$. This yields a collection of cells over which the fibers are empty. Note that $\Gr^{\tilde Q}_\tbfg(\tilde P_{m}^{\uj})$ is just a point in this case.

If $\tbfg\neq \udim\tilde P_{m+1}^{(k)}$, by Proposition~\ref{fibers}, the fiber dimension only depends on the dimension vectors $\tbff, \tbfg$ which means that they are constant over the cells of $\Gr^{\tilde Q}_\tbff(P_{m-1,j_l})\times \Gr^{\tilde Q}_\tbfg(\tilde P_{m}^{\uj})$. In particular, they are constant over cells whose subrepresentations $U\in \Gr^{\tilde Q}_\tbff(P_{m-1,i_k})$ satisfy $\pi_m^{\bf i}(U)\neq 0$, but also over those whose subrepresentations $U\in \Gr^{\tilde Q}_\tbff(P_{m-1,i_k})$ satisfy $\pi_m^{\bf i}(U)= 0$. By Theorem~\ref{vb}, it follows that $\Gr^{\tilde Q}_\tbfe(\tilde P_{m}^{\uj'})$ has a cell decomposition.

Thus it remains to show that this cell decomposition of $\Gr^{\tilde Q}_\tbfe(\tilde P_{m}^{\uj'})$ is compatible with $\pi_{m,\uj'}^{\bf j}$ where ${\bf j}=(\underline{j_1},\underline{j_2},\ldots,\underline{j_r})$ with $\underline{j_1}=(\uj',j_{1,l},\ldots,j_{1,n-1})\in I_{n-1}$ and $j_{1,l}=j_l$. Recall that we have $\pi_{m,\uj'}^{\bf j}=\pi_{m,\uj}^{\bf j}\circ\pi_{m,\uj'}^{j_l}$ by Lemma~\ref{le:truncated quotients}.(2).

@@

By Lemma~\ref{directsums} and induction hypothesis every quiver Grassmannian on the right hand side has a cell decomposition. The second part of the Proposition says that the fibers are empty if and only if $\tbfg=0$ and $\tbfh=\udim \tilde P_{m+1}^{(k)}$. In all other cases, the fiber dimension only depend on the dimension vectors $\tbff, \tbfg,\tbfh$ which means that they are constant over $\Gr^{\tilde Q}_\tbff( P(m,k))\times \Gr^{\tilde Q}_\tbfg(\tau\tilde P_{m+1}^{(k)})\times \Gr^{\tilde Q}_\tbfh(\tilde P_{m+1}^{(k)})$\footnote{actually this should already say that this is a vector bundle as the base space is smooth (all reps are exceptional), apply Lemma~\ref{vb}?}. Analogously to Lemma~\ref{directsums}, we can conclude that this induces trivial bundles over every affine cell. In total, this yields a cell decomposition of the quiver Grassmannian $\Gr^{\tilde Q}_\tbfe(\tilde P_{m+1}^{(k-1)})$.
\end{comment}
\end{proof}

@@

\begin{theorem}\label{celldec}The following holds:
\begin{enumerate}
\item Every quiver Grassmannian of any indecomposable preprojective or preinjective representation of $K(n)$ and $\widetilde{K(n)}$ has a cell decomposition into affine spaces.
\item Let $X\in\rep K(n)$ be an indecomposable representation of dimension $(d,e)$ or $(e,d)$ where $(d,e)=\udim P_m-r\udim P_{m-1}$ for $m\geq 2$ and $\leq 0\leq r\leq n-1$. Then every quiver Grassmannian $\Gr_{\bfe}(X)$ has a cell decomposition into affine spaces.
\end{enumerate}
\end{theorem}
\begin{proof}
The first part follows by combining the results of Section~\ref{torusaction} and Theorems~\ref{thm:bb} and~\ref{cellscover} taking into account that every preprojective representation of $\widetilde{ K(n)}$ is a lift of a preprojective representation which can be proven using the same argument as in the proof of Lemma~\ref{le:lifts of transjectives}. The dual version for preinjective representations follows as $\Gr_{\bfe}(P_m)\cong \Gr_{\udim P_m-\bfe}(I_m)$ and $\Gr_{\tbfe}(\tilde P_m)\cong\Gr_{\udim \tilde P_m-\tbfe}(\tilde I_m)$ respectively.

The second part follows in the same way, additionally, taking Theorem~\ref{thm:truncpp} into account.
\end{proof}

@@

By Theorem~\ref{celldec}, the claim is true for all quiver Grassmannians attached to $P_{j_i}$ or $I_{k_i}$. Consider the short exact sequence
\[\ses{P_{j_{r'+1}}}{P(r'+1)}{P(r')}.\]
By induction, we can assume that all quiver Grassmannians attached to the two outer terms have a cell decomposition into affine spaces. Consider the Caldero-Chapoton map
\[\Psi_\bfe:\Gr_\bfe(P(r'+1))\to\bigsqcup_{\bff+\bfg=\bfe} \Gr_\bff(P_{j_{r'+1}})\times \Gr_\bfg(P(r')).\]
The results of \cite[Section 3]{cc} show that $\Psi_{\tbfe}^{-1}(U,V)=\AA^{\dim\Hom(V,P_{j_{r'+1}}/U)}$  for all $(U,V)\in \Gr_\bff(P_{j_{r'+1}})\times \Gr_\bfg(P(r'))$, in particular the fibre is never empty. Now every subrepresentation $V$ of $P(r')$ is isomorphic to direct sum of preprojective representations such that for each direct summand $P_l$ we have $l\leq j_{r'}$. Moreover, the quotient $P_{j_{r'+1}}/U$ is not projective if $U\neq 0$ and equal to $P_{j_{r'+1}}$ otherwise. Together this yields $\Ext(V,P_{j_{r'+1}}/U)=0$ and thus
$$\dim\Hom(V,P_{j_{r'+1}}/U)=\langle V, P_{j_{r'+1}}/U\rangle$$
for all $(U,V)\in \Gr_\bff(P_{j_{r'+1}})\times \Gr_\bfg(P(r'))$. By Theorem~\ref{vb}, this already shows that $\Gr_{\bfe}(P(r'))$ has a cell decomposition into affine spaces for every $1\leq r'\leq r$. By duality, the same is true for $\Gr_{\bfe}(I(s))$.

@@

As the $F$-polynomial of truncated preprojective representations only depend on the dimension vector, we may denote them by $F_{\udim P_m{(k)}}$. The description of the non-empty fibers in Proposition~\ref{fibers} together with Corollary~\ref{fpoly} and Theorem~\ref{celldec} yield the following:
\begin{corollary}
For $m\geq 2$ and $0\leq k\leq n-2$, we have 
$$F_{\udim P_m^{(k)}}=F_{\udim P_m^{(k+1)}}F_{\udim P_{m-1}}-x^{\udim P_{m}^{(k)}}F_{\udim P_{m-2}^{(k)}}.$$
\end{corollary}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Combinatorial description of non-empty cells}
\noindent In this section, we provide two combinatorial descriptions of the non-empty cells in the quiver Grassmannians of (truncated) preprojective representations of $K(n)$.
The first is quiver theoretic and follows directly from the recursive construction of the cell decomposition in Section~\ref{sec:fibrations}.
The second is the notion of compatible pairs in a maximal Dyck path arising in the computation of rank 2 cluster variables \cite{llz}.
We give a bijection between these which provides a partial answer to a question of the first author \cite{rupel}.

@@

%The aim is to associate a $2$-quiver with every preprojective representation $P_m$ and with their quotients $P_m^{(k)}$ defined in section~\ref{QG}.
%These $2$-quivers are related to certain coefficient quivers of these representations.
%The upshot is that the non-empty cells of $\Gr_e(P_m)$ and $\Gr_e(P_m^{(k)})$ are in one-to-one correspondence with the strong successor closed subsets $\beta$ of type $e$.
%Hence the corresponding Euler characteristic can be computed when counting these subsets.
%It will turn out that a feature of this description is that it does not take care of the coloring of the different arrows induced by the arrows of $K(n)$.

%\begin{remark}\footnote{Insert a remark that we do not work on the universal cover} 
%  Lemma~\ref{AR} can be generalized to the cases $m\leq 2$.
%  The only technical problem is that $\tau \tilde P^{(k)}_{m+1}$ is no quotient of $\tilde P_{m,k}$ and we have to replace it by the image of the unique morphism to $\tau \tilde P^{(k)}_{m+1}$ which induces the exact sequence. 
%\end{remark}
In the following, we frequently use the notation and conventions of Section~\ref{Lifting}.
The next step is to associate a $2$-quiver $\cQ_{m+1}^{[k]}$ with the lifted truncated preprojective representations 
\[\tilde P_{m+1}^{[k]}:=
  \begin{cases}
    \tilde P_{m+1}^{\{n,n-1,\ldots,n-k+1\}} & \text{ if $m$ is odd;}\\
    \tilde P_{m+1}^{\{1,2,\ldots,k\}} & \text{ if $m$ is even;}
  \end{cases}\] 
for $m\geq 1$ and $0\leq k\leq n-1$.
When $k=0$ above, we take $\tilde P_{m+1}^{[0]}=\tilde P_{m+1}$.
We aim to construct $\cQ_{m+1}^{[k]}$ so that the strong successor closed subsets are in one-to-one correspondence with cells of quiver Grassmannians of $\tilde P_{m+1}^{[k]}$.

@@

\item  For $m\geq 3$, recall that we have $\dim\Hom(P_{m-1},\tau P_m^{[k]})=\dim\Ext(P_m^{[k]},P_{m-1})$ by the Auslander-Reiten formula. This induces isomorphisms
\begin{align}
  \label{eq:homext} 
  \bigoplus_{i=k+1}^n\Hom(\alpha_{i}^{(-1)^{m}}.\tilde P_{m-1},\tau\tilde P_{m}^{[k]})\cong\bigoplus_{i=k+1}^n\Ext(\tilde P_m^{[k]},\alpha_{i}^{(-1)^{m}}. \tilde P_{m-1})\cong \Ext(P_m^{[k]},P_{m-1}).
\end{align}
This means that we can choose a basis of $\Ext(P_m^{[k]},P_{m-1})$ which is compatible with the lifts to the universal covering and which gives a one-to-one correspondence between the basis elements of $\Ext(P_m^{[k]},P_{m-1})$ and $\Hom(P_{m-1},\tau P_m^{[k]})$. 
In view of Corollary~\ref{cor:perpendicular}, we have also an isomorphism $$\Ext(\tilde P_m^{[k]},\tilde P_{m-1,k})\cong\Ext(\tilde P_m^{[k]},\tau\tilde P_m^{[k]}).$$
These isomorphisms will also be present in the structure of the $2$-quivers $\cQ_m^{[k]}$. Finally note that there exists a short exact sequence
\[\ses{\Hom(P_{m-1},P_{m-1})}{\Ext(P_m^{[k]},P_{m-1})}{\Ext(P_m^{[k+1]},P_{m-1})}.\]
Thus a basis of $\Ext(P_m^{[k]},P_{m-1})$ can be obtained by taking the last $k$ basis elements of a basis of $\Ext(P_m^{[n-1]},P_{m-1})$.
\end{enumerate}
\end{remark}
